---
title: "Citation Information"
author: "Abigail Stamm"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Citation Information}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Citation and feedback

Suggestions for improvement and bug reports are welcome. Send feedback to [NYSDOH EPHT](mailto:epht@health.ny.gov?subject=[Geomask in R]) and include *Geomask in R* in the subject line.

If you use any part of the Geomasking Tool (Geomasker) package, please cite it. For the suggested citation, type:

```{r}
citation(package = "geomask")
```

## Acknowledgements

The conversion of the Geomasker to an R package would not have been possible without the support of the Centers for Disease Control and Prevention (CDC) National Environmental Health Tracking Program (NEPHT) and NYS Department of Health (NYSDOH) Bureau of Environmental and Occupational Epidemiology (BEOE) staff and interns, including:

**Major contributors**  
Abigail Stamm, NYSDOH BEOE  
Gwen (Babcock) LaSelva, NYSDOH BEOE  

**Alpha testers**  

**Beta testers**  

The Geomasker was made possible through funding made available to the New York State Department of Health / Health Research, Incorporated by the CDC through grant # NUE1EH001341-03-01. Its content is solely the responsibility of the authors and does not necessarily represent the official views of the CDC.

